---
title: "README for consmap-prep"
author: "Ben Best"
date: "October 6, 2015"
output: html_document
---

# Overview

Once you have species distributions, how do you use them to make decisions? We're encouraged through multiple avenues (Presidential Order, Pew Report...) to move away from single species management to a holistic ecosystem-based management. So we want to combine the species distributions, but how?

Borrowing from [Maxwell et al (2013)](http://www.nature.com.proxy.lib.duke.edu/ncomms/2013/131028/ncomms3688/full/ncomms3688.html), we can apply species-specific impact weight. For instance, more acoustically sensitive species have a stronger weight based on offshore pile driving as a human activity.

TODO: explore weighting, sensitivity analysis:

- industry weights
- literature explore weighting:
  - multi-criteria decision analysis
  - IUCN /  NatureServe advice on weighting
  - [Selig et al (2013)](http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0060284)
- diagram of inputs / process

Questions:
1. Rather than divide by guild, should I divide by species, ie multiply guilds by # of species?

# Industry


- wind intensity
- distance to shore
- depth
- slope

- distance to shore / transmission. with/without Atlantic Wind Connection

```{r load libraries, echo=FALSE}
suppressPackageStartupMessages({
  library(raster)
  library(ggmap)
  library(leaflet)
  library(RColorBrewer)
  library(dplyr)
  library(readr)
  library(tidyr)
  library(DT)
  library(xts)
  library(dygraphs)
  library(graticule)
  library(rphylopic)
  #library(GISTools) # using GISTools::
  })
select = dplyr::select

source('functions.R')
stk_i = stack('data/industry_wind_mer.grd')
```

```{r interactive-plots}
# TODO: condense into single stack, opacity solid.
iplot(raster(stk_i, 'depth'), 'depth')
iplot(raster(stk_i, 'dist2shore'), 'dist2shore')
iplot(raster(stk_i, 'windspeed90m'), 'windspeed90m')
```

# Species

## Species Weights

The species weights (**w**) are based on the NatureServe global extinction status which have a reasonable value for all species, unlike IUCN for which many are data deficient (DD).

```{r spp, echo=FALSE}
# noa pbr table ----

# exclude Harbor seal, Gray seal Harp seal Hooded seal
# all excepting 'Common bottlenose dolphin': 
#   EC:
#     Gulf of Maine 
#     Western North Atlantic 
#     Nova Scotia 
#     Canadian east coast 
#     North Atlantic 
#     Gulf of Maine/Bay of Fundy 
#   GOM:
#     Gulf of Mexico
#     Puerto Rico and US Virgin Islands
# 'Common bottlenose dolphin' (n=28): 
#   EC: 
#     Western North Atlantic, offshore 
#     Western North Atlantic, northern migratory coastal 
#     Western North Atlantic, southern migratory coastal 
#     Western North Atlantic, S. Carolina/Georgia coastal 
#     Western North Atlantic, northern Florida coastal 
#     Western North Atlantic, central Florida coastal 
#     Northern North Carolina Estuarine System 
#     Southern North Carolina Estuarine System 
#     Northern South Carolina Estuarine System 
#     Charleston Estuarine System 
#     Northern Georgia/ Southern South Carolina Estuarine System 
#     Central Georgia Estuarine System 
#     Southern Georgia Estuarine System 
#     Jacksonville Estuarine System 
#     Indian River Lagoon Estuarine System 
#     Biscayne Bay 
#   GOM:
#     Florida Bay 
#     Gulf of Mexico, Continental shelf 
#     Gulf of Mexico, eastern coastal 
#     Gulf of Mexico, northern coastal 
#     Gulf of Mexico, western coastal 
#     Gulf of Mexico, Oceanic 
#     Gulf of Mexico, bay, sound and estuary (27 stocks) 
#     Barataria Bay 
#     Mississippi Sound, Lake Borgne, Bay Boudreau 
#     St. Joseph Bay 
#     Choctawhatchee Bay 
#     Puerto Rico and US Virgin Islands 
#  - two morphologically and genetically distinct common bottlenose dolphin morphotypes described as the 
#    **coastal** and **offshore** forms. spatially varies by season, and overlaps.

# strip annotations/whitespaces and convert to numeric
d_sar = read_csv('data/species/spp_noaa-marmam_atl2014.csv') %>%
    mutate(
      species    = str_trim(species),
      stock_area = str_trim(stock_area),
      nmfs_ctr   = str_trim(nmfs_ctr),
      N_best     = as.numeric(str_replace_all(N_best,    '[^0-9.]', '')),
      N_best_CV  = as.numeric(str_replace_all(N_best_CV, '[^0-9.]', '')),
      N_min      = as.numeric(str_replace_all(N_min,     '[^0-9.]', '')),
      R_max      = as.numeric(str_replace_all(R_max,     '[^0-9.]', '')),
      Fr         = as.numeric(str_replace_all(Fr,        '[^0-9.]', '')),
      PBR        = as.numeric(str_replace_all(PBR,       '[^0-9.]', '')),
      injury_mortality          = as.numeric(ifelse(
        # take upper estimate, eg '5.8' given '3.8-5.8'
        str_detect(injury_mortality_fish_0,'-'),
        str_replace(injury_mortality_fish_0,'.*-([0-9.]+).*','\\1'),
        str_replace_all(injury_mortality, '[^0-9.]', ''))),
      injury_mortality_fish     = as.numeric(ifelse(
        # take upper estimate, eg '0.8' given '0.2-0.8'
        str_detect(injury_mortality_fish_0,'-'),
        str_replace(injury_mortality_fish_0,'.*-([0-9.]+).*','\\1'),
        ifelse(
          # take value if given CV, eg '3.4' given '3.4 (1.0)'
          str_detect(injury_mortality_fish_0,'\\('),
          str_replace(injury_mortality_fish_0,'([0-9.]+)[ (].*','\\1'),
          str_replace_all(injury_mortality_fish_0, '[^0-9.]', '')))),
      injury_mortality_fish_CV  = as.numeric(ifelse(
        # take CV if given, eg '1.0' given '3.4 (1.0)'
        str_detect(injury_mortality_fish_0,'\\('),
        str_replace(injury_mortality_fish_0,'.*\\(([0-9.]+)\\)','\\1'),
        NA)),
      injury_mortality_ship = as.numeric(injury_mortality_ship),
      strategic_status = str_trim(strategic_status),
      SAR_revised = str_trim(SAR_revised))

# combine all EC species (1 stock ea for EC) and summarize Common bottlenose dolphins to single stock
d_pbr = bind_rows(
  # EC species, except Common bottlenose dolphin and seals
  d_sar %>%
    filter(
      !species %in% c(
        'Common bottlenose dolphin',
        'Harbor seal','Gray seal','Harp seal','Hooded seal'),
      !stock_area %in% c('Gulf of Mexico','Puerto Rico and US Virgin Islands')),
  # Brydes using Gulf of Mexico since none given for EC
  #   from East Coast Bryde's Whale v1.2.pdf:
  #     South of Gulf Stream: Estimated abundance: 7 (CV=0.58)
  #     North of Gulf Stream: Species assumed absent; forced model to zero.)
  d_sar %>% # arrange(species) %>% datatable()
    filter(species == "Bryde's whale"),
  d_sar %>%
    filter(
      !species %in% c(
        'Common bottlenose dolphin',
        'Harbor seal','Gray seal','Harp seal','Hooded seal'),
      !stock_area %in% c('Gulf of Mexico','Puerto Rico and US Virgin Islands')),
  # EC Common bottlenose dolphin stocks summarized to single stock
  d_sar %>%
    filter(
      species == 'Common bottlenose dolphin',
      stock_area %in% c(
        'Western North Atlantic, offshore','Western North Atlantic, northern migratory coastal',
        'Western North Atlantic, southern migratory coastal','Western North Atlantic, S. Carolina/Georgia coastal',
        'Western North Atlantic, northern Florida coastal','Western North Atlantic, central Florida coastal',
        'Northern North Carolina Estuarine System','Southern North Carolina Estuarine System',
        'Northern South Carolina Estuarine System','Charleston Estuarine System',
        'Northern Georgia/ Southern South Carolina Estuarine System','Central Georgia Estuarine System',
        'Southern Georgia Estuarine System','Jacksonville Estuarine System','Indian River Lagoon Estuarine System',
        'Biscayne Bay')) %>%
    group_by(species) %>%
    summarize(
      stock_area = 'Combined North Atlantic',
      nmfs_ctr   = paste(unique(nmfs_ctr), collapse=','),
      N_best_CV  = weighted.mean(N_best_CV, N_best, na.rm=T),
      R_max      = weighted.mean(R_max, N_best, na.rm=T),
      Fr         = weighted.mean(Fr, N_best, na.rm=T),
      injury_mortality_fish_CV  = weighted.mean(injury_mortality_fish_CV, N_best, na.rm=T),
      # note: have to use weighted.mean before sum(N_best) b/c data frame modified in place by dplyr
      N_best     = sum(N_best, na.rm=T), # ) %>% datatable() # 
      N_min      = sum(N_min, na.rm=T),
      PBR        = sum(PBR, na.rm=T),
      injury_mortality          = sum(injury_mortality, na.rm=T),
      injury_mortality_fish     = sum(injury_mortality_fish, na.rm=T),
      injury_mortality_ship     = sum(injury_mortality_ship, na.rm=T),
      strategic_status          = 'Combined',
      SAR_revised               = 'Combined'))
d_pbr %>% datatable()

spp = read_csv('data/species/spp.csv') %>%
  left_join(read_csv('data/species/spp_weights.csv'), by='natureserve') %>%
  merge(d_pbr, by.x='sp_name', by.y='species', all.x=T)
datatable(spp) # , options = list(pageLength = 5)

w_spp = spp %>% 
  group_by(sp_code) %>%
  summarize(w = mean(w)) %>%
  arrange(desc(w))
datatable(w_spp)

paste(intersect(spp$sp_name, d_pbr$species), collapse="','")
#paste(setdiff(spp$sp_name, d_pbr$species), collapse="','")
cat(paste(setdiff(d_pbr$species, spp$sp_name), collapse=": \n"))
# Dwarf sperm whale: Kogia sima
# Pygmy sperm whale: Kogia breviceps
# Cuvier's beaked whale: Ziphius cavirostris
# Blainville's beaked whale: Mesoplodon densirostris
# Gervais beaked whale: Mesoplodon europaeus
# Sowerby's beaked whale: Mesoplodon bidens
# True's beaked whale: Mesoplodon mirus
# Melon-headed whale: Peponocephala electra
# Pilot whale, long-finned: Globicephala melas
# Pilot whale, short-finned: Globicephala macrorhynchus
# White-beaked dolphin: Lagenorhynchus albirostris
# Common bottlenose dolphin: Tursiops truncatus

```

## Birds

```{r birds tables prep, eval=F, echo=F}
b_lyr = read_csv('data/birds/sp_AvianAverageAnnualAbundance.csv') %>%
  select(-sp_scientific)
  arrange(sp_common)
b_pbr = read_csv('data/birds/watts2010_atlantic-birds_pbr.csv') %>%
  mutate(
    sp_common    = str_trim(sp_common),
    species_popn = str_trim(species_popn)) %>%
  arrange(sp_common, species_popn)
b_popn = read_csv('data/birds/watts2010_atlantic-birds_popn.csv') %>%
  mutate(
    sp_common = str_trim(sp_common),
    species_popn = str_trim(species_popn)) %>%
  arrange(sp_common, species_popn)

b = b_lyr %>%
  left_join(
    b_pbr %>%
      full_join(
        b_popn, by=c('sp_common','species_popn','aou')),
    by='sp_common')
# head(b)
# datatable(b)
write_csv(b, 'data/birds/sp_avian_watts2010.csv')

b %>%
  transmute(
    sp_common,
    has_pbr = !is.na(pbr),
    has_popn_na = !is.na(popn_na)) %>%
  select(-sp_common) %>%
  table() %>% addmargins()
# 
#        has_popn_na
# has_pbr FALSE TRUE Sum
#   FALSE     1   20  21
#   TRUE      7    6  13
#   Sum       8   26  34
  
# make bradbury [OLD]
read_csv('data/birds/bradbury2014_uk-bird-sensitivity-scores.csv') %>%
  mutate(
    sp_common    = str_trim(sp_common),
    sp_scientific = str_trim(sp_scientific)) %>%
  left_join(
    read_csv('data/birds/bradbury2014_uk-bird-sensitivity-risk-ranks.csv') %>%
      mutate(sp_common = str_trim(sp_common))
  , by='sp_common') %>%
  write_csv('data/birds/bradbury2014_uk-bird-sensitivity-scores_2.csv')
  datatable()
  

b = b %>% 
  mutate(
    sp_scientific = str_replace(species_popn, '([A-z]+) ([A-z]+).*', '\\1 \\2'))
b %>% select(starts_with('sp')) %>% 
  group_by(sp_common, sp_scientific) %>%
  summarize(sp_popns_watts2010 = n()) %>%
  mutate(sp_scientific = replace(sp_scientific, sp_common=='Great Shearwater', 'Puffinus gravis')) %>%
  write_csv('data/birds/sp_avianshp_scientific.csv')

b = read_csv('data/birds/sp_avianshp_scientific.csv')
b %>%
  left_join(u, by='sp_scientific') %>%
  select(starts_with('sp'), cons_tot) %>%
  filter(is.na(cons_tot)) %>%
  # datatable()
  write_csv('data/birds/sp_avianshp_scientific_bradbury2014unmatched_notes.csv')
  

# calculating Bradbury 2014
# a = score for highest percent of biogeographic population in England in any season; 
#   TODO: adjust above to US Atlantic
# b = adult survival score;
# c = UK threat status score;
#   TODO: adjust above to US Atlantic
# d = Birds Directive score;
# total conservation imp score ----
# e = estimated percentage at blade height; 
# f = flight manoeuvrability;
# g = percentage of time spent flying;
# h = nocturnal activity;
# i = disturbance susceptibility;
# j = habitat specialization.
bu = 
b %>%
  left_join(u, by='sp_scientific') %>%
  filter(!is.na(cons_tot)) %>%
  #datatable()
  mutate(
    # collision risk score
    cons_calc = a + b + c + d,
    cons_test = cons_calc == cons_tot,
    # collision risk score = a x (m + t + n)/3 
    # [flight altitude] x ([manoeuvrability] + [percentage of time spent flying] + [nocturnal flight activity])/3
    coll_risk_score = e * (f + g + h)/3,
    coll_risk_rank = cut(coll_risk_score, c(30,50,82), labels=c('Very low','Low','Moderate','High','Very high'), include.lowest=T, ordered_result=T),
    # displacement score = ((d x h) x conservation importance score)/10
    # (([disturbance by wind farm structures, ship and helicopter traffic] x [habitat specialisation]) x conservation importance score)/10
    displ_score = ((i * j) * cons_calc)/10
    ) %>%
  select(sp_common.y, e, f, g, h, coll_risk_score, coll_risk_rank, cons_calc, displ_score) %>%
  datatable()
# TODO: contact Bradbury about mismatch b/n scores

table(!is.na(b$pbr), !is.na(b$popn_na))

paste(intersect(b_lyr$sp_common, b_pbr$sp_common), collapse="','")
paste(setdiff(b_lyr$sp_common, b_pbr$sp_common), collapse="','")
paste(setdiff(b_pbr$sp_common, b_lyr$sp_common), collapse="','")
cat(paste(setdiff(d_pbr$species, spp$sp_name), collapse=": \n"))


b = read_csv('data/birds/sp_avian.csv')
b %>%
  left_join(
    read_csv('data/birds/bradbury2014_uk-bird-sensitivity-scores.csv'), 
    by='sp_scientific') %>%
  filter(!is.na(coll_rank)) %>%
  mutate(
    coll_rank  = str_trim(coll_rank),
    displ_rank = str_trim(displ_rank),
    coll_val  = c('Very low'=1,'Low'=2,'Moderate'=3,'High'=4,'Very high'=5)[coll_rank],
    displ_val = c('Very low'=1,'Low'=2,'Moderate'=3,'High'=4)[displ_rank]) %>% 
  select(
    sp_scientific, sp_common_us=sp_common.x, sp_common_uk=sp_common.y,
    coll_rank, coll_val, displ_rank, displ_val) %>%
  arrange(sp_scientific) %>%
  write_csv('data/birds/sp_avian_w.csv')
```


```{r birds shps, eval=F, echo=F}

b_w = read_csv('data/birds/sp_avian_w.csv') %>%
  mutate(
    shp = str_replace_all(sp_common_us, "[ '-]",''))

dir_in = 'data_other/msp/AvianAverageAnnualAbundance/shp'
dir_out = 'data_other/msp/AvianAverageAnnualAbundance/shp'
x = 0

if (exists('stk_b')) rm('stk_b')

r_i = raster('data/wind/npv_US_millions_8_idwf.tif')
stk_c = stack('data/species/spp_EC_nzw_aea.grd')
r_c = raster(stk_c, 'ALL_nfzw') # iplot(r_C,'C_nfzw')

for (shp in list.files(dir_in, pattern='.*\\.shp$')){ # shp = list.files(dir_shp, pattern='.*\\.shp$')[1]
  
  pfx = tools::file_path_sans_ext(shp)
  s_c     = b_w %>% filter(str_to_lower(b_w$shp) == str_to_lower(pfx)) %>% .$sp_scientific %>% str_replace(' ', '_')
  
  w_displ = b_w %>% filter(str_to_lower(b_w$shp) == str_to_lower(pfx)) %>% .$displ_val
  w_coll  = b_w %>% filter(str_to_lower(b_w$shp) == str_to_lower(pfx)) %>% .$coll_val
  
  if (length(w_coll)==0){
    cat(sprintf('%s: SKIPPING b/c no w_coll\n', shp))
    next # skip if no w
  } else {
    w = max(w_coll, w_displ)
    cat(sprintf('%s: w_displ=%d, w_coll=%d, w=%d\n', shp, w_displ, w_coll, w))
  }
  
  ply = readOGR(dir_in, tools::file_path_sans_ext(shp), verbose=F, p4s=leaflet:::epsg3857) %>%
    spTransform(crs(r_c))
  ply@data = ply@data %>%
    mutate(
      avg_min = as.numeric(ifelse(
        str_detect(averageAnn,' - '),
        str_replace(averageAnn, '([0-9.]+) - ([0-9.]+)', '\\1'),
        str_replace(averageAnn, '([0-9.]+) \\+', '\\1'))),
      avg_max = as.numeric(ifelse(
        str_detect(averageAnn,' - '),
        str_replace(averageAnn, '([0-9.]+) - ([0-9.]+)', '\\2'),
        str_replace(averageAnn, '([0-9.]+) \\+', '\\1'))),
      avg = (avg_min + avg_max) / 2
    ) # spplot(ply['avg'])
  r_n = rasterize(ply, r_c, 'avg') # iplot(r_n)
  r_nw = log(r_n + 1) * w # plot(r_nlwc)
  
  # n = abundance; nlwc = ln(n+1)*SSI_collision; nlwd = ln(n+1)*SSI_displacement; nlwm = ln(n+1)*SSI_max where SSI_max[collision or displacement]
  # TODO: n not real, based on strip transects details...
  nms = sprintf('%s_%s', s_c, c('n','nw'))
  
  if (!exists('stk_b')){
    stk_b = stack(r_n, r_nw)
    names(stk_b) = nms
  } else {
    stk_b = stack(stk_b, r_n, r_nw)
    names(stk_b) = c(names(stk_b)[1:(nlayers(stk_b)-length(nms))], nms)
  }
}

substk = function(x, v){ # x=stk; s='_nf'
  ext = sprintf('_%s', v)
  lyrs = names(x)[str_sub(names(x), nchar(ext)*-1, -1) == ext]
  raster::subset(x, lyrs)
}
    
# a_i: number of species
s_i = substk(stk_b, 'n')
a_i = sum(!is.na(s_i), na.rm=T) # a_i_0 = a_i
# reconcile edge effects by requiring all species present
a_i = mask(a_i, a_i==cellStats(a_i, 'max'), maskvalue=0) # iplot(a_i,'a_i')

# a_n: average species abundance
a_n = substk(stk_b, 'n') %>%
  sum(na.rm=T) / a_i %>%
  mask(a_i) # iplot(a_n)

# a_nw: average log species density * SSI_max weight
a_nw = substk(stk_b, 'nw') %>%
  sum(na.rm=T) / a_i %>%
  mask(a_i) # iplot(a_nw)

# TODO: contouring to make as pretty as Figure 6 of Bradbury 2014

# add to stack
stk_b = stack(stk_b, a_i, a_n, a_nw)
names(stk_b) = c(names(stk_b)[1:(nlayers(stk_b)-3)], sprintf('BIRDS_%s', c('i','n','nw')))

# write out raster stack per region
writeRaster(stk_b, 'data/birds/spp_birds_aea.grd', overwrite=T)

names(stk_b)
#v = 'Larus_argentatus_n'; iplot(raster(stk_b, v), v)
#v = 'Larus_argentatus_nw'; iplot(raster(stk_b, v), v)
#stk_b = stack('data/birds/sp_birds.grd')
```


## Birds Map

```{r prep birds ma r_depth}
# bird rasters
s_b = stack('data/birds/spp_birds_aea.grd')
r_b = raster(s_b, 'BIRDS_nw')

# crop to birds to extent
m <- is.na(as.matrix(r_b)) # %>% addmargins()
cols = which(colSums(m) != nrow(m))
rows = which(rowSums(m) != ncol(m))
e_c = extent(
  r_b, 
  rows[1], rows[length(rows)],
  cols[1], cols[length(cols)])
r_b <- crop(r_b, e_c) # plot(r_b)

# get USA states
us = getData('GADM', country="USA", level=1)
us@data = us@data %>%
  mutate(
    lon = coordinates(us)[,1],
    lat = coordinates(us)[,2],
    code = c(setNames(state.abb, state.name), 'District of Columbia'='DC')[NAME_1])
# get East Coast states
ec = subset(us,
  code %in% 
    c('ME','VT', # NE
      'NY','MA','RI','CT','PA','NJ','MD','DE','DC','WV','VA','NC', # Mid-Atlantic
      'SC','GA','FL')) # South
# ec@data %>%
#   arrange(desc(pt_lat)) %>%
#   .$code %>% paste0("','") %>% cat(sep='')
# get Mid-Atlantic states
ma_gcs = subset(us,
  code %in% 
    c('NY','MA','RI','CT','PA','NJ','MD','DE','DC','WV','VA','NC')) # Mid-Atlantic
# project to aea
ma = spTransform(ma_gcs, crs(r_b))
# get plotting coordinates
ma@data = ma@data %>%
  mutate(
    x = coordinates(ma)[,1],
    y = coordinates(ma)[,2])
ma_pts_manual = list( # get from plotting raster and using click()
  NY=c(318279.0, 861861.0),
  MA=c(583509.0, 900670.0),
  PA=c(176331.0, 753784.4),
  NJ=c(302205.0, 697670.0),
  DE=c(218600.4, 532496.3),
  MD=c(116658.7, 604601.4),
  VA=c(87289.93, 403240.4), #58407.0, 403619.0),
  NC=c(142765.7, 194348.2))
for (code in names(ma_pts_manual)){
  xy = ma_pts_manual[[code]]
  ma@data$x[ma@data$code==code] = xy[1]
  ma@data$y[ma@data$code==code] = xy[2]
}

library(marmap)
r_depth = as.raster(getNOAA.bathy(-78, -66, 34, 43))
```


```{r map birds}
plot_raster(r_b, r_depth, ma, pdf_path='img/study_mid-atlantic.pdf', awc=T, study_area=T, par_mar=c(4,4,0,0)+0.2) # c(bottom, left, top, right)
plot_raster(r_b, r_depth, ma, pdf_path='img/birds_mid-atlantic.pdf', phylopic='herring gull', depth_do = F, quantile_do = T)
```


## InVEST + Birds

```{r map invest}
r_i6 =  mask(projectRaster(raster('data/wind/npv_US_millions_6.tif'), r_b), r_b)
r_i8 =  mask(projectRaster(raster('data/wind/npv_US_millions_8_idwf.tif'), r_b), r_b)

plot_raster(r_i6, r_depth, ma, pdf_path='img/industry_npv6_mid-atlantic.pdf', phylopic='wind turbine')
plot_raster(r_i8, r_depth, ma, pdf_path='img/industry_npv8idwf_mid-atlantic.pdf', awc=T, phylopic='wind turbine')
```

```{r map cetaceans}
s_c  = stack('data/species/spp_EC_nzw_aea.grd')
r_c = raster(s_c,'ALL_nfzw') %>%
  mask(extend(r_b, r_c)) %>%
  crop_na() #; iplot(r_c, 'nzw')

r_c = r_c
plot_raster(r_c, r_depth, ma, pdf_path='img/cetaceans_mid-atlantic.pdf', phylopic='balaena')

for (i in 1:12){
  r_c_i = raster(s_c,sprintf('ALL_nfzw_%02d', i)) %>%
    mask(extend(r_b, .)) %>%
    crop_na() #; iplot(r_c, sprintf('nzw_c%02d', i))
  plot_raster(r_c_i, r_depth, ma, pdf_path=sprintf('img/cetaceans_mid-atlantic_%02d.pdf',i), phylopic='balaena')
}
```


Scale 0-1.

```{r birds vs industry iplot, eval=F, echo=F}
library(ggplot2)
library(RColorBrewer)
library(scales)
library(sp)
library(rgeos)

r_b = raster(stack('data/birds/spp_birds_aea.grd'), 'BIRDS_nw') %>%
  crop_na() #; iplot(r_b, 'r_b')
r_d = raster(stack('data/industry_wind_aea.grd'), 'depth') %>%
  crop(r_b) %>%
  mask(r_b) %>%
  crop_na() #; iplot(r_d, 'r_d')
r_i = raster('data/wind/npv_US_millions_8_idwf.tif') %>%
  projectRaster(r_b) %>% 
  crop(r_b) %>%
  mask(r_b) %>%
  crop_na() #; iplot(r_i, 'r_i')

# get birds vs invest plot and identify 
S = stack(r_i, r_b)
d = values(S) %>%
  as.data.frame() %>%
  select(
    B_ssi = BIRDS_nw, 
    I_usm = npv_US_millions_8_idwf) %>%
  mutate(
    i = row_number()) %>%
  filter(!is.na(B_ssi) & !is.na(I_usm)) %>%
  # consider only profitable sites, ie y > 0
  filter(I_usm > 0) %>%
  # normalize 0 to 1
  mutate(
    x = rescale(B_ssi),
    y = rescale(I_usm))

# test plot
#par(mar=c(5,4,1,1)) # c(bottom, left, top, right)
#with(d, plot(x, y, xlab='B', ylab='I', xlim=rev(range(x))))
#di = identify(d$x, x$y, col='red', labels=x$i, cex=0.8, font=2)
#dput(di) # xi = c(3L, 23L, 24L, 25L, 569L, 719L, 770L, 820L, 845L, 1539L, 1566L, 1570L)
#points(d$x[di], d$y[di], col='red')

# get efficiency frontier with convex hull of points
i_h = chull(d$x, d$y) # lines(d$x[i_h], d$y[i_h], col='blue'); text(d$x[i_h], d$y[i_h], d$i[i_h], col='red')
# then abline to max(y), abline to max(x)
h = data_frame(
  x = d$x[i_h],
  y = d$y[i_h])
d_f = bind_rows(
  # bottom, right
  data_frame(
    x = min(h$x),
    y = min(h$y)),
  # frontier curve
  filter(h,
    y >= y[which.min(x)],
    x <= x[which.max(y)]),
  # top, left
  data_frame(
    x = max(h$x),
    y = max(h$y))) %>%
  mutate(
    i = row_number())
#lines(d_f$x, d_f$y, col='red')
#text(d_f$x, d_f$y, d_f$i, col='blue')

# get indices for frontier
i_f = semi_join(d, filter(h,
    y >= y[which.min(x)],
    x <= x[which.max(y)]), by=c('x','y')) %>%
  .$i  # 4256 4081 4138 1615  918
lbl_f = LETTERS[1:length(i_f)]

# to get distance to frontier ----
# - make polygon of convex hull
# - make grid points
# - calculate distance from frontier line to grid points

# make poly for within
ply = SpatialPolygons(list(Polygons(list(Polygon(bind_rows(
  # bottom, left
  data_frame(
    x = max(h$x),
    y = min(h$y)),
  # frontier curve
  d_f %>% select(x, y),
  # bottom, left
  data_frame(
    x = max(h$x),
    y = min(h$y))) %>%
  select(x, y) %>%
  as.matrix())), 1)))
# expand ply by 1 to include frontier edge
ply = rgeos::gBuffer(ply, width=0.01) #plot(ply, add=T, col='gray')

# make grid of points
z_n = 100
pts_grid = expand.grid(x = seq(min(d_f$x), max(d_f$x), length.out = z_n),
            y = seq(min(d_f$y), max(d_f$y), length.out = z_n)) %>%
  SpatialPoints() # %>% points(col='purple')
# remove points outside poly
pts_grid = pts_grid[!is.na(over(pts_grid, ply))] # %>% points(col='black')

# calculate distance to frontier for plotting
ln = SpatialLines(list(Lines(list(Line(d_f %>% select(x, y) %>% as.matrix())), '1')))
g = coordinates(pts_grid) %>% as.data.frame()
g$z = as.vector(gDistance(pts_grid, ln, byid=TRUE))
#g$z_log = log(g$z)
#g$z_log = log10(g$z)
g$z_log = g$z

# plot 
brks = c(0,0.25,0.5,0.75,1)
lbls_B = sprintf('%0.3f', approx(x=d$x, y=d$B_ssi, xout=brks, rule=c(2,2))$y)
lbls_I = sprintf('%0.1f', approx(x=d$y, y=d$I_usm, xout=brks, rule=c(2,2))$y)

p = ggplot(d, aes(x, y)) +
  #geom_raster(aes(x, y, z = z, fill=z), data=g) +  #, size=1, color=z
  geom_raster(aes(x, y, z = z_log, fill=z_log), data=g) +  #, size=1, color=z
  #geom_tile(aes(x, y, z = z_log, fill=z_log), data=g) +  #, size=1, color=z
  scale_fill_distiller(name='', palette='Spectral', trans='reverse', breaks = pretty_breaks(n = 10)) +
  #geom_contour(aes(x, y, z = z, fill=z), data=g, color = "white", alpha = 0.5) + 
  geom_contour(aes(x, y, z = z_log, fill=z_log), data=g, color = "white", alpha = 0.5) + 
  #geom_path(data=select(d_f, x, y), col=brewer.pal(11, 'Spectral')[10], lwd=6) + 
  geom_point() + 
  geom_text(aes(x, y, label=lbl_f), data = d[match(i_f, d$i),], colour='gray90', size=4, hjust=1, vjust=1) +
  geom_point(aes(x, y), data = d[match(i_f, d$i),], colour='gray90', size=1) + 
  scale_x_continuous(
     name = 'Bird Sensitivity', breaks=brks, labels=lbls_B, trans='reverse') +
   scale_y_continuous(
     name = 'Industry NPV', breaks=brks, labels=lbls_I) +
  coord_cartesian(xlim = c(-0.01,1), ylim = c(0,1.01)) +
  theme_bw()+
  theme(
    legend.title = element_blank())
p
pdf = 'img/birds_vs_industry8.pdf'
ggsave(pdf, p, scale=0.5)
system(sprintf('open %s', pdf))

# calculate distance to frontier for raster ----
d$d2frontier = as.vector(gDistance(SpatialPoints(d[,c('x','y')]), ln, byid=TRUE))

r_f = r_b
r_f[!1:ncell(r_f) %in% d$i] = 0
r_f_l10 = r_f_l = r_f
r_f[d$i] = d$d2frontier
r_f_l[d$i] = log(d$d2frontier + 0.000001) # %>% range ; iplot(r_f)
r_f_l10[d$i] = log(d$d2frontier + 0.000001) # %>% range ; iplot(r_f)
r_f[d$i] = log10(d$d2frontier + 0.000001) # %>% range
r_f = mask(r_f, r_b); iplot(r_f)

ll_pts = SpatialPoints(coordinates(r_b)[i_f,], proj4string=crs(r_b)) %>%
      spTransform(crs(leaflet:::epsg4326)) %>%
      coordinates()
lbl_pts = data_frame(
  i     = i_f,
  label = lbl_f,
  x     = coordinates(r_b)[i_f,'x'],
  y     = coordinates(r_b)[i_f,'y'],
  lon   = ll_pts[,'x'],
  lat   = ll_pts[,'y'],
  B_ssi = r_b[i_f],
  I_usm = r_i[i_f],
  d2frontier_log = d$d2frontier[match(i_f, d$i)],
  d2frontier_log = log(d$d2frontier[match(i_f, d$i)]) + 0.000001,
  d2frontier_log10 = log10(d$d2frontier[match(i_f, d$i)]) + 0.000001); lbl_pts

plot_raster(r_f, r_depth, ma, pdf_path='img/map_birds_vs_industry8_frontier_log.pdf', lbl_pts = lbl_pts, lbl_col='white', awc=T)
plot_raster(r_f, r_depth, ma, pdf_path='img/map_birds_vs_industry8_frontier_log.pdf', lbl_pts = lbl_pts, lbl_col='white', awc=T)
write_csv(lbl_pts, 'data/birds_vs_industry8_frontierpts.csv')
```

```{r cetaceans ma time, echo=FALSE}
#lbl_pts = read_csv('data/birds_vs_industry8_frontierpts.csv')

# now with utility hotspots (vs dist2frontier)
lbl_pts = read_csv(
'label,i
E,1615
B,4138
F,918
D,2095
C,3430
A,4322')

r_b = raster(stack('data/birds/spp_birds_aea.grd'), 'BIRDS_nw') %>%
  crop_na() #; iplot(r_b, 'r_b')
s_c = stack('data/species/spp_EC_nzw_aea.grd') %>% # plot(raster(s_c,'ALL_nfzw'))
  crop(r_b) %>%
  mask(r_b)
# r_c = raster(s_c,'ALL_nfzw') %>% # plot(raster(s_c,'ALL_nfzw'))
#   crop(r_b) %>%
#   mask(r_b) %>%
#   crop_na() #; plot(r_b) plot(r_c), 'nzw')
r_c = raster(s_c,'ALL_nfzw') %>% # plot(raster(s_c,'ALL_nfzw'))
  crop_na() #; plot(r_b) plot(r_c), 'nzw')

#v = data_frame()
# c_ll@data %>%
#   select(id, val) %>%
#   merge(data_frame(mo = 1:12))
suppressWarnings(rm('v'))
for (mo in 1:12){ # mo = 1
  # get raster
  lyr = sprintf('ALL_nfzw_%02d', mo)
  r = raster(s_c, lyr) %>%
    crop_na() 
  #plot(r, main=lyr)
  # TODO: plot some of these cetacean sensitivity maps by lowest months (3,5,8,11,12) with site labels
  
  # extract values
  v_mo = data.frame(
    label = lbl_pts$label,
    mo    = mo,
    val   = r[lbl_pts$i])

  # bind to stack
  if (!exists('v')){
    v = v_mo
  } else {
    v = bind_rows(v, v_mo)
  }
}

summary(v$val)
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# -10.200  -5.460  -0.193  -1.460   1.990   5.410 

# rescale values 0 to 1
v = v %>%
  rename(site=label) %>%
  mutate(
    val = rescale(val))
summary(v$val)
#  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 0.000   0.305   0.642   0.561   0.781   1.000

# TODO: redo conservation risk index so doesn't go negative, outputs are decimal, and rescaled from quantiles across entire year of population estimates.

# get month per site with minimum value
v_min = v %>% 
  group_by(site) %>% 
  mutate(site_rank = rank(val)) %>%
  arrange(site, site_rank, val) %>%
  filter(site_rank == 1)

img_whale = image_data('18438875-f6e8-4e46-8fda-5a07bd4c1a85', size=1024)[[1]]
f_pdf = 'img/cetaceans_time_utilitypts_phylopicwhale.pdf'
pdf(f_pdf)
p = ggplot(v, aes(x=mo, y=val, col=site)) + 
  geom_line(size=1) +
  coord_cartesian(xlim = c(1, 12)) + 
  xlab('Month') + ylab('Cetacean Sensitivity') + # scale_x_date(format = "%b-%Y") + 
  #scale_x_continuous(breaks=c(1,3,6,9,12), minor_breaks=seq(1,12,by=1), labels=month.abb[c(1,3,6,9,12)]) +
  scale_x_discrete(labels=month.abb) +
  geom_point(aes(x=mo, y=val, col=site), data=v_min, cex=5) + 
  theme(legend.justification=c(1,1), legend.position=c(1,1)) + 
  add_phylopic(img_whale, alpha=1, x=9.3, y=0.98, ysize=1)
print(p)
dev.off()
system(sprintf('open %s', f_pdf))
#ggsave('img/cetaceans_time_frontierpts_phylopicwhale.png')
```

```{r map cetacean sensitivity}
lbl_pts = pts %>%
  left_join(
    d_sum, by='i') %>%
  rename(x_0=x, y_0=y) %>%
  inner_join(
    coordinates(r_b) %>%
      as.data.frame() %>%
      mutate(
        i = row_number()),
    by='i')

# TODO: plot some of these cetacean sensitivity maps by lowest months (3,5,8,11,12) with site labels
source('functions.R')
#for (mo in 1:12){ # mo = 3
for (mo in c(3,5,8,11,12)){ # mo = 3
  r_c = raster(s_c, sprintf('ALL_nfzw_%02d', mo))
  # TODO: fix this rescale hack like mentioned in above TODO:
  #   redo conservation risk index so doesn't go negative, outputs are decimal, 
  #   and rescaled from quantiles across entire year of population estimates.
  values(r_c) = rescale(values(r_c))
  if (mo==3){
    plot_raster(
      r_c, r_depth, ma, pdf_path=sprintf('img/map_cetacean-sensitivity_mo%02d-%s_phylopic.pdf', mo, month.abb[mo]), 
      awc=T, depth_do = F, quantile_do = F, lbl_pts = lbl_pts, lbl_col='black', phylopic='balaena')
  } else {
    plot_raster(
      r_c, r_depth, ma, pdf_path=sprintf('img/map_cetacean-sensitivity_mo%02d-%s.pdf', mo, month.abb[mo]), 
      awc=T, depth_do = F, quantile_do = F, lbl_pts = lbl_pts, lbl_col='black')
  }
}
  
  

lyr = 
r =  %>%
  crop_na() 
#plot(r, main=lyr)




```



```{r old dygraph}
# old interactive ----    
# TODO: animation of graph changing over time
mos = as.Date(sprintf('2015-%02d-01', 1:12))
v %>%
  spread(site, val) %>%
  replace(is.na(.), 0) %>%
  arrange(mo) %>%
  select(-mo) %>%
  as.xts(., order.by=mos, names=months(mos, abbreviate=T)) %>%
  dygraph() %>%
  #dyHighlight() %>%
  dyHighlight(highlightSeriesBackgroundAlpha = 0.2) #%>%
    #highlightCircleSize = 5, 
    #hideOnMouseOut = FALSE) %>%
  #dyOptions(stepPlot=T) 

# TODO: show avg per cell
# %>% dySeries("V1", strokeWidth = 2, strokePattern = "dashed")

# TODO: upper/lower bars w/ sdm confidence
#   http://rstudio.github.io/dygraphs/gallery-upper-lower-bars.html
```




```{r tradeoff plot birds vs industry}
r_B  = raster(S, 'BIRDS_nw')
r_I  = raster(S, 'npv_US_millions_8_idwf')


# what is x? this chunk presumably a fragment of above
c_id  = x$i[xi]
c_n   = length(c_id)
c_xy  = coordinates(S)[c_id,]
c_pts = SpatialPointsDataFrame(
  c_xy, proj4string=CRS(proj4string(S)), data=data.frame(i=1:c_n)) #,
c_ll = spTransform(c_pts, leaflet:::epsg4326)
c_ll@data = c_ll@data %>%
  mutate(
    id  = c_id,
    lbl = c_id,
    B   = r_B[c_id],
    I   = r_I[c_id],
    #col = brewer.pal(c_n, 'Spectral'),
    x   = coordinates(c_xy)[,'x'],
    y   = coordinates(c_xy)[,'y'],
    lon = coordinates(c_ll)[,'x'],
    lat = coordinates(c_ll)[,'y'])
with(c_ll@data, plot(B, I, xlim=rev(range(B))))

# TODO: check in ArcGIS since pts are off grids.
writeRaster(r_i, 'data/birds/r_i_8_idwf_aea.tif')
writeRaster(r_b, 'data/birds/r_b_nw_aea.tif')
writeOGR(c_pts, 'data/birds', 'birds_vs_industry_pts_aea', driver='ESRI Shapefile')
writeOGR(c_ll, 'data/birds', 'birds_vs_industry_pts_gcs', driver='ESRI Shapefile')

r_I_mer = projectRasterForLeaflet(r_I)  
r_B_mer = projectRasterForLeaflet(r_B)  
r = r_B_mer # r_I_mer r_I
r = r_I_mer # r_I_mer r_I
# get color palette
pal = colorNumeric(
  rev(brewer.pal(11, 'Spectral')), 
  values(r), na.color = "transparent")

# interactive map
leaflet() %>%
  addTiles() %>%
  addProviderTiles("Stamen.TonerLite", options = providerTileOptions(noWrap = TRUE)) %>%
  addRasterImage(r, colors=pal, project=F, opacity=0.8, group='r') %>% # CAREFUL! assumes r = projectRasterForLeaflet(r_0)  
  addLegend(pal=pal, position='bottomright', values=values(r), title='r') %>%
  #addMarkers(data=c_ll, popup = ~value)
  addMarkers(
    data=c_ll,
    popup = sprintf('<b>%d.<b><br>B: %0.3f<br>I: %0.3f', c_ll$lbl, c_ll$B, c_ll$I), group='Points') %>%
  addLayersControl(
    overlayGroups = c('r', 'Points'),
    options = layersControlOptions(collapsed=T))

iplot(r_i8)
s_ec_mer  = stack('data/species/spp_EC_nzw_mer.grd')
```

## EC

```{r quickfix, eval=F, echo=F}
for (prj in c('mer','aea')){ # prj='mer'
  
  cat(sprintf('reading grd_%s\n', prj))
  s_0     = stack(sprintf('data/species/spp_EC_nzw_%s_0.grd', prj))
  s_1to12 = stack(sprintf('data/species/spp_EC_nzw_%s_1to12.grd', prj))
  s = stack(s_0, s_1to12)
  
  cat(sprintf('writing grd_%s\n', prj))
  writeRaster(s, sprintf('data/species/spp_EC_nzw_%s.grd', prj), overwrite=F)
}
```

```{r cetaceans_ec, echo=FALSE}
s_ec_mer  = stack('data/species/spp_EC_nzw_mer.grd')

iplot(raster(s_ec_mer,'Brydes_whale_n'), 'Brydes_whale_n')
hist(values(raster(s_ec_mer,'Brydes_whale_n')))
table(values(raster(s_ec_mer,'Brydes_whale_n')))
names(s_ec_mer)
iplot(raster(s_ec_mer,'ALL_i'), 'i')
iplot(raster(s_ec_mer,'ALL_nf'), 'n')
iplot(raster(s_ec_mer,'ALL_nfz'), 'nz')
iplot(raster(s_ec_mer,'ALL_nfzw'), 'nzw')
```

: avg density [n; sum(n)/i; # / 100km^2]
: avg weighted normalized density (nzw = nz * w)

## EC, Over Time

```{r cetaceans_ec_time, echo=FALSE}
r = raster(s_ec_mer,'ALL_nfzw')

# cells 
# manually get c_id's by clicking on map:  plot(r); click(r, cell=T, id=T)
c_id  = c(41935, 16016, 12146, 30162, 34222)
c_n   = length(c_id)
c_xy  = coordinates(r)[c_id,]
c_pts = SpatialPointsDataFrame(
  c_xy, proj4string=CRS(proj4string(r)), data=data.frame(i=1:c_n)) #,
c_ll = spTransform(c_pts, leaflet:::epsg4326)

c_ll@data = c_ll@data %>%
  mutate(
    id  = c_id,
    val = r[c_id],
    col = brewer.pal(c_n, 'Spectral'),
    x   = coordinates(c_xy)[,'x'],
    y   = coordinates(c_xy)[,'y'],
    lon = coordinates(c_ll)[,'x'],
    lat = coordinates(c_ll)[,'y'])

# get color palette
pal = colorNumeric(
  rev(brewer.pal(11, 'Spectral')), 
  values(r), na.color = "transparent")

# interactive map
leaflet() %>%
  addTiles() %>%
  addProviderTiles("Stamen.TonerLite", options = providerTileOptions(noWrap = TRUE)) %>%
  addRasterImage(r, colors=pal, project=F, opacity=0.8, group='Raster nfzw') %>%
  addLegend(pal=pal, position='bottomright', values=values(r), title='nfzw') %>%
  #addMarkers(data=c_ll, popup = ~value)
  addMarkers(
    data=c_ll,
    popup = sprintf('%d. nfzw: %0.1f', c_ll$i, c_ll$val), group='Points') %>%
  addLayersControl(
    overlayGroups = c('Raster nfzw', 'Points'),
    options = layersControlOptions(collapsed=T))


#v = data_frame()
# c_ll@data %>%
#   select(id, val) %>%
#   merge(data_frame(mo = 1:12))
suppressWarnings(rm('v'))
for (mo in 1:12){ # mo = 6
  # get raster
  r = raster(s_ec_mer, sprintf('ALL_nfzw_%02d', mo))
  #plot(r)
  
  # extract values
  v_mo = data.frame(
    i   = 1:c_n,
    mo  = mo,
    val = r[c_id])

  # bind to stack
  if (!exists('v')){
    v = v_mo
  } else {
    v = bind_rows(v, v_mo)
  }
}

plot(r)
text(c_xy, as.character(1:c_n))

# TODO: animation of graph changing over time
mos = as.Date(sprintf('2015-%02d-01', 1:12))
v %>%
  spread(i, val) %>%
  replace(is.na(.), 0) %>%
  arrange(mo) %>%
  select(-mo) %>%
  as.xts(., order.by=mos, names=months(mos, abbreviate=T)) %>%
  dygraph() %>%
  #dyHighlight() %>%
  dyHighlight(highlightSeriesBackgroundAlpha = 0.2) %>%
    #highlightCircleSize = 5, 
    #hideOnMouseOut = FALSE) %>%
  dyOptions(stepPlot=T) 

# TODO: show avg per cell
# %>% dySeries("V1", strokeWidth = 2, strokePattern = "dashed")

# TODO: upper/lower bars w/ sdm confidence
#   http://rstudio.github.io/dygraphs/gallery-upper-lower-bars.html
```

TODO: zoom into pixel and break down across time by species.

## GOM

```{r cetaceans_gom, echo=FALSE}
s_gom_mer = stack('data/species/spp_GOM_nzw_mer.grd')

iplot(raster(s_gom_mer,'ALL_i'), 'i')
iplot(raster(s_gom_mer,'ALL_nf'), 'n')
iplot(raster(s_gom_mer,'ALL_nfz'), 'nz')
iplot(raster(s_gom_mer,'ALL_nfzw'), 'nzw')
```


```{r create_old, eval=FALSE, echo=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(leaflet)
  library(raster)})

source('functions.R')
r_depth = raster('data/physiographic/EC_Depth_10km_mean_10km.img')
plot(raster('data/physiographic/EC_Depth_10km_mean_10km.img'))
plot(raster('data/physiographic/EC_Shore_Dist_10km_mean_10km.img'))
plot(raster('data/physiographic/EC_Slope_10km_mean_10km.img'))

library(rgdal)
windspeed90m = readOGR(
  'data/wind_intensity/Atlantic_Coast_90m_Windspeed_Offshore_Wind_High_Resolution',
  'atlantic_coast_90mwindspeed_off')
plot(windspeed90m)
hist(windspeed90m@data$Speed_90) # 5-10 m/s

windspeed90m_aea = spTransform(windspeed90m, CRS(proj4string(r_depth)))
r_windspeed90m_aea = rasterize(windspeed90m_aea, r_depth, 'Speed_90')
writeRaster(r_windspeed90m_aea, 'data/wind/windspeed90m_aea.grd')
iplot(r_windspeed90m_aea)
```

