---
title: "utility_simulation"
author: "Ben Best"
date: "January 27, 2016"
output: html_document
---

## Utility Simulation

Inspired by chat with Chris Costello Jan 12, 2016.

$$
u = \alpha N + (1-\alpha)C
$$

where:

- $u$: utility
- $\alpha$: tradeoff coefficient
- $N$: net present value to industry
- $C$: conservation of species

questions:

- how is utility tangent to efficiency frontier, but equation above has no concept of that tangent?

```{r get rasters}
# setwd('~/github/consmap-prep')
suppressPackageStartupMessages({
  library(raster)
  library(ggmap)
  library(leaflet)
  library(RColorBrewer)
  library(dplyr)
  library(readr)
  library(tidyr)
  library(DT)
  library(xts)
  library(dygraphs)
  library(graticule)
  library(rphylopic)
  #library(GISTools) # using GISTools::
  library(scales)
  })
select = dplyr::select
source('functions.R')
source('utility_simulation_app/functions.R')

r_b = raster(stack('data/birds/spp_birds_aea.grd'), 'BIRDS_nw') %>%
  crop_na() #; iplot(r_b, 'r_b')
r_d = raster(stack('data/industry_wind_aea.grd'), 'depth') %>%
  crop(r_b) %>%
  mask(r_b) %>%
  crop_na() #; iplot(r_d, 'r_d')
r_i = raster('data/wind/npv_US_millions_8_idwf.tif') %>%
  projectRaster(r_b) %>% 
  crop(r_b) %>%
  mask(r_b) %>%
  crop_na() #; iplot(r_i, 'r_i')
```

Garthe & Hüppop (2004) designated the top 60% of scores as areas of “concern” and top 20% as areas of “major concern”. 

```{r contour birds}
q_b20 = quantile(r_b, probs=0.2)
q_b60 = quantile(r_b, probs=0.6)
plot(r_b)
contour(
  r_b, add=TRUE, drawlabels=F,
  levels = c(q_b20, q_b60), 
  #labels = c(                   '20%',                    '60%'), 
  col    = c(                   'red',                   'blue'))
```


```{r tradeoff data}
# get birds vs invest plot and identify 
S = stack(r_i, r_b)
d = values(S) %>%
  as.data.frame() %>%
  select(
    B_ssi = BIRDS_nw, 
    I_usm = npv_US_millions_8_idwf) %>%
  mutate(
    i = row_number()) %>%
  filter(!is.na(B_ssi) & !is.na(I_usm)) %>%
  # consider only profitable sites, ie y > 0
  filter(I_usm > 0) %>%
  # normalize 0 to 1
  mutate(
    x = rescale(B_ssi),
    y = rescale(I_usm))

# get pareto rows
d = d %>%
  mutate(pareto = is_pareto(x, y, rev_x=T))

# get pareto labels
d = d %>%
  left_join(
    d %>% 
      filter(pareto) %>%
      arrange(x) %>%
      mutate(
        label = LETTERS[row_number()],
        alpha = (1 - y) / (x - y)) %>%
      select(i, label, alpha),
    by='i') %>%
  arrange(pareto, label, desc(y), desc(x)) %>%
  tbl_df()

save(d, file='utility_simulation_app/d.Rdata')
```

```{r plot pareto with dots}
# ggplot birds vs industry
#ggplot(d, aes(B_ssi, I_usm)) +
#  geom_point()
#ggplot(d, aes(x, y)) +
#  geom_point()

# plot with pareto dots
load(file='utility_simulation_app/d.Rdata')
#xlim=rev(range(B)))
ggplot(d, aes(x, y)) +
  geom_point() +
  scale_x_continuous(trans='reverse') +
  geom_point(data=filter(d, pareto), color='red') +
  geom_text(aes(x, y, label=label), data = filter(d, pareto), colour='red', size=4, hjust=1, vjust=1)
```

```{r table of pareto points}
# show pareto points
d %>%
  filter(pareto) %>%
  datatable()
```


```{r create raster surfaces, eval=F}
# create x,y,u rasters
library(raster)
library(rasterVis)
n = 110
x = seq(0,1.1,length.out=n)
y = seq(0,1.1,length.out=n)
z = sqrt(outer(x^2, y^2, "+"))
r_x = raster(matrix(x,nrow=n,ncol=n, byrow=T), xmn=min(x), xmx=max(x), ymn=min(y), ymx=max(y)) # plot(r_x)
r_y = raster(matrix(y,nrow=n,ncol=n, byrow=F), xmn=min(x), xmx=max(x), ymn=min(y), ymx=max(y)) %>%
  flip('y') # plot(r_y)
#r_z = raster(z, xmn=min(x), xmx=max(x), ymn=min(y), ymx=max(y)) %>%
#  flip('y') # plot(r_z)
#r_z = sqrt(r_x^2 + r_y^2)
#plot(r_z)
#levelplot(sqrt(r_x^2 + r_y^2), par.settings=RdBuTheme, contour=T) # rasterVis::levelplot
```


```{r manipulate ggplot, eval=F}
# manipulate ggplot
library(manipulate)
manipulate({ # a = 0
  
  # calculate utility surface
  r_u = a * r_x + (1 - a) * r_y # plot(r_z)

  # calculate utility for points
  d = d %>%
    mutate(
      u = a * x + (1 - a) * y,
      u_max = u == max(u))
  
  # get line, per y = mx + b
  u_slope         = - a/(1 - a)
  u_max_intercept = filter(d, u_max)$u / (1 - a)

  # # TODO: plot with pareto dots colored by value
  # ggplot(d, aes(x, y, color=u)) +
  #   #geom_point(color='gray') +
  #   geom_point() +
  #   #scale_colour_gradient(limits=c(3, 4), low="red")
  #   #scale_colour_brewer(type='seq', palette='YlOrRd') + # RColorBrewer::display.brewer.all()
  #   #scale_colour_distiller(name='', palette='Spectral', trans='reverse', breaks = pretty_breaks(n = 10)) +
  #   scale_colour_distiller(name='', palette='Spectral', breaks = pretty_breaks(n = 10)) +
  #   geom_point(data=filter(d, pareto), color='black') +
  #   geom_point(data=filter(d, u_max), color='red') +
  #   geom_abline(slope = u_slope, intercept = u_max_intercept, color='red') +
  #   geom_text(aes(x, y, label=label), data = filter(d, pareto), colour='black', size=4, hjust=1, vjust=1)
    
  # plot raster
  p = rasterVis::gplot(r_u, maxpixels=n^2+1) + 
    geom_tile(aes(fill = value)) +
    scale_fill_distiller(name='', palette='Spectral', trans='reverse', breaks = pretty_breaks(n = 10)) +
    #geom_contour(aes(x=x,y=y,fill=value), color = "white", alpha = 0.5) + 
    coord_equal() + 
    scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
    geom_point(aes(x, y), data=d, color='gray') + 
    geom_point(aes(x, y), data=filter(d, pareto), color='black') +
    geom_text(aes(x, y, label=label), data = filter(d, pareto), colour='black', size=4, hjust=1, vjust=1) +
    ggtitle(sprintf('a = %0.2f, %s', a, filter(d, u_max)$label)) +
    geom_vline(xintercept = quantile(d$x, probs=0.8, na.rm=T), linetype=1, col='red') +
    geom_vline(xintercept = quantile(d$x, probs=0.4, na.rm=T), linetype=1, col='blue')
  
  # so which a?
  # for siting owed, want to max conservation, so place owed at site w/ least bird cons value. -> flip axes.
  # label: bird conservation value 1 to 0. "Sensitivity to birds"
  

  if (is.infinite(u_slope)){
    # when u_slope is Inf (a=1), draw vline
    x_intercept = filter(d, u_max)$x
    p = p +
      geom_vline(xintercept = x_intercept)
  } else {
    p = p +
      geom_abline(slope = u_slope, intercept = u_max_intercept)
  }
  print(p)
  
}, a = slider(0, 1, step=0.05))
```

rank sites across range of reasonable values
TODO: compare ranks when using original values vs normalized


```{r sim full}
# simulate across range of values
d_sim = simulate_utility(
  d, x='B_ssi', y='I_usm', a_vec = seq(0, 1, 0.1), 
  fxn=lazyeval::lazy( a * y - (1 - a) * x ),
  fxn_slope=lazyeval::lazy( -1*(1 - a)/a ), 
  fxn_intercept=lazyeval::lazy( u[u_max] / a ))

# summarize sites by sum of utility across alphas
d_sum = d_sim %>%
  group_by(i, x, y) %>%
  summarize(
    u_avg = mean(u_sim)) %>%
  ungroup() %>%
  arrange(desc(u_avg)) %>%
  mutate(
    rank = row_number(),
    pct_rank = percent_rank(u_avg)) #%>%
  # left_join(
  #   d_sim %>%
  #     filter(u_max==T) %>%
  #     arrange(a_sim) %>%
  #     group_by(i) %>%
  #     mutate(
  #       a_sim_m = median(a_sim)
  #     )
  #     summarize(
  #       a_sim   = first(a_sim),
  #       u_slope = first(u_slope),
  #       u_yintercept = first(u_yintercept)) %>%
  #     mutate(
  #       u_max = T) %>%
  #     select(i, a_sim, u_max, u_slope, u_yintercept),
  #   by='i') %>%
  # mutate(
  #   u_max = ifelse(is.na(u_max), F, T))
write_csv(d_sum, '~/github/siting/data/utility/utility_birds-vs-industry8_v2_data.csv')

# print pareto sites
# d_sum %>%
#   left_join(
#     filter(d, pareto) %>% 
#       tbl_df() %>%
#       select(i, label), 
#     by='i') %>%
#   arrange(desc(u_avg)) %>%
#   datatable()
```


```{r plot tradeoff of simulation}
# get ticks for x and y based on original values
brks = c(0,0.25,0.5,0.75,1)
lbls_B = sprintf('%0.3f', approx(x=d$x, y=d$B_ssi, xout=brks, rule=c(2,2))$y)
lbls_I = sprintf('%0.1f', approx(x=d$y, y=d$I_usm, xout=brks, rule=c(2,2))$y)

d_a = d_sim %>%
  filter(u_max==T) %>%
  arrange(a_sim)

pts = read_csv(
'label,i
E,1615
B,4138
F,918
D,2095
C,3430
A,4322')

# summary plot
sz = 4 # text size
p = ggplot(d_sum, aes(x, y, colour=u_avg)) +
  # rectangles for quadrant descriptions
  annotate('rect', fill='blue', alpha=0.1, xmin=0, ymax=1,
    xmax=quantile(d_sum$x, probs=0.4, na.rm=T),
    ymin=quantile(d_sum$y, probs=0.6, na.rm=T)) +
  annotate('rect', fill='green', alpha=0.1,
    xmin=0,
    xmax=quantile(d_sum$x, probs=0.8, na.rm=T),
    ymin=quantile(d_sum$y, probs=0.2, na.rm=T),
    ymax=quantile(d_sum$y, probs=0.6, na.rm=T)) +
  annotate('rect', fill='green', alpha=0.1,
    xmin=quantile(d_sum$x, probs=0.4, na.rm=T),
    xmax=quantile(d_sum$x, probs=0.8, na.rm=T),
    ymin=quantile(d_sum$y, probs=0.6, na.rm=T),
    ymax=1) +
  annotate('rect', fill='red', alpha=0.1,
    xmin=0,
    xmax=1,
    ymin=0,
    ymax=quantile(d_sum$y, probs=0.2, na.rm=T)) +
  annotate('rect', fill='red', alpha=0.1,
    xmin=quantile(d_sum$x, probs=0.8, na.rm=T),
    xmax=1,
    ymin=quantile(d_sum$y, probs=0.2, na.rm=T),
    ymax=1) +
  # points
  geom_point() +
  #xlim(0,1) + ylim(0,1) +
  #coord_fixed(ratio = 1, xlim=c(0,1), ylim=c(0,1), expand=F) +
  coord_equal(xlim=c(0,1), ylim=c(0,1), expand=F) +
  #coord_cartesian(xlim=c(0,1), ylim=c(0,1), expand=F) +
  scale_x_continuous(
     name = 'Bird Sensitivity', breaks=brks, labels=lbls_B, trans='reverse') +
   scale_y_continuous(
     name = 'Wind Profitablity ($NPV)', breaks=brks, labels=lbls_I) +
  scale_colour_gradientn(colours = brewer.pal(11, 'Spectral'), name=expression(bar(u))) + # brewer.pal(9, 'Blues') # YlGnBu # brewer.pal(9, 'GnBu'); brewer.pal(11, 'Spectral'); 
  with(filter(d_a, a_sim == median(a_sim)), geom_abline(slope = u_slope, intercept = u_yintercept, linetype=3)) + # median a
  # with(filter(d_sum, a_sim == min(a_sim, na.rm=T)),                                                                          # min a
  #   {if ( is.infinite(u_slope) ) geom_vline(                                     
  #     xintercept = x, linetype=3) else geom_abline(
  #       slope = u_slope, intercept = u_yintercept, linetype=3)}) +
  # with(filter(d_sum, a_sim == max(a_sim, na.rm=T)),                                                                          # max a
  #   {if ( is.infinite(u_slope) ) geom_vline(                                     
  #     xintercept = x, linetype=3) else geom_abline(
  #       slope = u_slope, intercept = u_yintercept, linetype=3)}) +
  geom_vline(xintercept = quantile(d_sum$x, probs=(1-0.2), na.rm=T), linetype=2, col='red') +
  geom_vline(xintercept = quantile(d_sum$x, probs=(1-0.6), na.rm=T), linetype=2, col='blue') +
  geom_hline(yintercept = quantile(d_sum$y, probs=   0.2 , na.rm=T), linetype=2, col='red') +
  geom_hline(yintercept = quantile(d_sum$y, probs=   0.6 , na.rm=T), linetype=2, col='blue') +
  # annotations, see: 
  #   https://github.com/hadley/ggplot2/wiki/Plotmath
  #   http://vis.supstat.com/2013/04/mathematical-annotation-in-r/
  annotate('text', x = 0.08, y = .88, label = 'tilde(a)', parse=T, size=sz, angle=-45) +
  annotate(
    'text', label = "' ' %up% 20~'%'~most~sensitive", parse=T, size=sz, angle=90, vjust=0, hjust=0,
    x = quantile(d_sum$x, probs=(1-0.2), na.rm=T), 
    y = quantile(d_sum$y, probs=0.01, na.rm=T)) +
  annotate(
    'text', label = "' ' %up% 60~'%'", parse=T, size=sz, angle=90, vjust=0, hjust=0,
    x = quantile(d_sum$x, probs=(1-0.6), na.rm=T), 
    y = quantile(d_sum$y, probs=0.01, na.rm=T)) +
  annotate(
    'text', label = "' ' %down% 20~'%'~least~profitable", parse=T, size=sz, vjust=1, hjust=0,
    x = quantile(d_sum$x, probs=0.99, na.rm=T), 
    y = quantile(d_sum$y, probs=0.2, na.rm=T)) +
  annotate(
    'text', label = "' ' %down% 60~'%'", parse=T, size=sz, vjust=1, hjust=0,
    x = quantile(d_sum$x, probs=0.99, na.rm=T), 
    y = quantile(d_sum$y, probs=0.6, na.rm=T)) +
  geom_text(
    aes(x, y, label=label), colour='black', size=4, hjust=0, vjust=1, data = pts %>%
      left_join(d_sum, by='i')) +
  geom_point(
    aes(x, y, label=label), colour='black', size=0.3, data = pts %>%
      left_join(d_sum, by='i'))

  # theme(
  #   panel.background = element_rect(fill='gray95'),
  #   panel.grid.major = element_line(colour = 'gray30'),
  #   panel.grid.minor = element_line(colour = 'gray30'))

# library(plotly)
# ggplotly(p)

pdf = 'img/birds_vs_industry8_v2-utility-simulation.pdf'
ggsave(pdf, p) # , scale=0.5)
system(sprintf('open %s', pdf))
```


```{r mk raster of utility}
# transfer dataframe to raster of avg utility
r_u = r_b
r_u[!1:ncell(r_u) %in% d_sum$i] = NA #-1 # min(d_sum$u_avg,na.rm=T) -> -0.2959718
r_u[d_sum$i] = d_sum$u_avg
r_u = mask(r_u, r_b) # iplot(r_u)
writeRaster(r_u, '~/github/siting/data/utility/utility_birds-vs-industry8_v2_raster')

# raster of ids
r_i = r_u
values(r_i) = 1:ncell(r_i)
```


```{r click raster, eval=F}
col = rev(colorRampPalette(brewer.pal(11,'Spectral'))(255))
plot(r_u, col=col, xaxs='r', yaxs='r', asp=1, useRaster=F, xaxt='n', yaxt='n')

quantile_vals = c(0.4,0.6,0.8)
quantile_cols = c('darkblue','darkgreen','darkred')
quantile_labs = c('60%','40%','20%')
quantile_lwd = 2
contour(r_u, levels = quantile(r_u, probs=quantile_vals), col=quantile_cols, labels=quantile_labs, drawlabels=T, lwd=quantile_lwd, add=T)

# interactively select corners of bbox
sub = raster::select(r_u)
d_sum %>%
  filter(i %in% raster::extract(r_i, sub@extent)) %>%
  arrange(desc(u_avg)) %>%
  select(i, rank) %>%
  head(1)

# 1615,1
# 4138,4
#  918,28
# 2095,36
# 3430,62
```


```{r plot raster}
lbl_pts = pts %>%
  left_join(
    d_sum, by='i') %>%
  rename(x_0=x, y_0=y) %>%
  inner_join(
    coordinates(r_b) %>%
      as.data.frame() %>%
      mutate(
        i = row_number()),
    by='i')

source('functions.R')
plot_raster(
  r_u, r_depth, ma, pdf_path='img/map_birds_vs_industry8_utility.pdf', awc=T, depth_do = F, quantile_do = T,
  lbl_pts = lbl_pts, lbl_col='black')
```

```{r write_csv birds_vs_wind8_utilitypts}
op = options()
options(scipen=0)
options(digits=0)

lbl_pts %>%
  # get lon, lat
  bind_cols(
    SpatialPoints(
      lbl_pts %>% select(x, y) %>% as.data.frame(), proj4string=crs(r_b)) %>%
      spTransform(crs('+proj=longlat +datum=WGS84')) %>%
      coordinates() %>%
      as.data.frame() %>%
      select(lon=x, lat=y)) %>%
  select(label, i, x_aea=x, y_aea=y, lon, lat, u_avg, rank) %>%
  # get original data
  left_join(d %>% select(-label), by='i') %>%
  mutate(
    lon = round(lon,2),
    lat = round(lat,2),
    B_ssi = round(B_ssi,2),
    I_usm = round(I_usm,1),
    u_avg = round(u_avg,3)) %>%
  select(
    Label=label, Longitude=lon, Latitude=lat, 
    `Bird Sensitivity`=B_ssi, `Wind Profitability (NPV $USM)`=I_usm,
    `Avg Utility`=u_avg, `Rank`=rank) %>%
  write_csv('data/birds_vs_wind8_utilitypts.csv')
```


# TODO: redo cetacean conservation

```{r}



```

